---
title: "p8105_hw1_ml4424"
author: "Maggie Li (ml4424)"
date: "9/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(tidyverse)
```

## Problem 1

```{r create df}
p1_df <- tibble(
  samp = rnorm(10),
  log_vec = samp > 0,
  char_vec = c("a", "b", "c", "d", "e","f", "g", "h", "i", "j"),
  fac_vec = factor(c("bananas", "oranges", "apples",
                    "bananas", "oranges", "apples",
                    "bananas", "oranges", "apples",
                    "bananas")))
p1_df

# you can take the mean of samp, log_vec (I am assuming since it is a vector of binary numeric values, T = 1 and F = 0 values) since these are continuous variables, but cannot take mean of characters or factors since these are categorical variables

```

```{r numeric coercion}
as.numeric(p1_df$log_vec)
as.numeric(p1_df$char_vec)
mean(as.numeric(p1_df$fac_vec))

# as.numeric works with logical and factor variables. it is because the logical vector consists of binary numeric values T=1 and F=0, and factor variables are categorical variables but when converted to numeric they get coerced to numbered levels. this helps to explain why mean works with logical vectors, but not factor variables until they are coerced to numeric.
```

```{r convert logical}
as.numeric(p1_df$log_vec) * p1_df$samp 
# this works because log_vec is easily coerced to numeric, T=1 and F=0

as.factor(p1_df$log_vec) * p1_df$samp
# T and F values get coded as factors; 2 levels

as.numeric(as.factor(p1_df$log_vec)) * p1_df$samp
# T gets coded as 2 and False as 1 for numeric, and can multiple it by another numeric
```

## Problem 2
```{r}
data("penguins", package = "palmerpenguins")
summary(penguins)
nrow(penguins %>% filter(sex == "female"))
levels(penguins$species)
summary(penguins$species)
nrow(penguins)
ncol(penguins)
mean(penguins$flipper_length_mm)
```

The names in the penguins dataset are 
`r names(penguins)`. 

The values for penguin species are `r summary(penguins$species)` for `r levels(penguins$species)`, respectively.

The values for penguin island are `r summary(penguins$island)` for `r levels(penguins$island)`, respectively.

The values for penguin bill length (mm) are `r summary(penguins$bill_length_mm)` for the minimum, 1st quantile, median, mean, 3rd quantile, and maximum, respectively.

The values for penguin bill depth (mm) are `r summary(penguins$bill_depth_mm)` for the minimum, 1st quantile, median, mean, 3rd quantile, and maximum, respectively.

The values for penguin flipper length (mm) are `r summary(penguins$flipper_length_mm)` for the minimum, 1st quantile, median, mean, 3rd quantile, and maximum, respectively.

The values for body mass in grams are `r summary(penguins$body_mass_g)` for the minimum, 1st quantile, median, mean, 3rd quantile, and maximum, respectively.

There are `r nrow(penguins %>% filter(sex == "female"))` females and `r nrow(penguins %>% filter(sex == "male"))` males in the penguins dataset.

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length is `r mean(penguins$flipper_length_mm)`.

### Scatterplot
